---
title: "Leaflet"
---

```{r setup, include=FALSE}

library(leaflet)
library(raster)
library(terra)
library(htmlwidgets)
library(htmltools)
library(leafem)
library(DT)
library(kableExtra)
library(leaflet.extras)
library(leaflet.extras2)

```

```{r layers, include=FALSE}

ntems2000 <- rast("C:/Users/melser/Documents/GitHub/rasters/CA_forest_VLCE2_2000/CA_forest_VLCE2_2000.tif")

ntems2020 <- rast("C:/Users/melser/Documents/GitHub/rasters/CA_forest_VLCE2_2020/CA_forest_VLCE2_2020.tif")

aoi <- vect("D:/VI.shp") %>% project(crs(ntems2020))

ntems2000 <- crop(ntems2000, aoi, mask = TRUE)
ntems2020 <- crop(ntems2020, aoi, mask = TRUE)

ntems2000 <- terra::as.factor(ntems2000)
ntems2020 <- terra::as.factor(ntems2020)

ntems2000 <- terra::project(ntems2000, "+init=EPSG:4326")
ntems2020 <- terra::project(ntems2020, "+init=EPSG:4326")

df <- data.frame(
  ID <- c(0, 20, 31, 33, 50, 80, 81, 100, 210, 220, 230),
  name <- c("Unclassified", "Water", "Snow", "Exposed/Barren", "Shrubs", "Wetland", "Wetland-Treed", "Herbs", "Coniferous", "Broadleaf", "Mixed Wood")
)

colnames(df) <- c("ID", "Names")

set.cats(ntems2000, value = df)
set.cats(ntems2020, value = df)

col <- c("#686868", "#3333FF", "#CCFFFF", "#996633", "#FFFF00",
         "#993399", "#9933CC", "#CCFF33", "#006600", "#00CC00",
         "#CC9900")

coldf <- data.frame(
  df[1],
  col)

coltab(ntems2000) <- coldf
coltab(ntems2020) <- coldf

```

**The Basics:**

Many software and coding packages can be used to create interactive maps. In this particular demo, we will show you an easy way to produce interactive maps in leaflet - a popular R package. Leaflet is intuitive, fast, and has many arguments that allow us to easily adjust the way in which we present our maps. In the example below, we used NTEMs land cover from 2020 to product a simple interactive map for Vancouver Island. In the code below, you will notice that we add in a basemap using the 'addProviderTiles' argument. You can browse for available leaflet basemaps here: <https://leaflet-extras.github.io/leaflet-providers/preview/>

```{r leaflet, include=TRUE, warning = FALSE}


l <- leaflet(width = 600, height = 600)%>%
  
  #lets use some base imagery from ESRI
  addProviderTiles("Esri.WorldImagery") %>%
  
  #now lets add our NTEMS land cover from 2020
  addRasterImage(ntems2020, 
                 #you can increase the allowable size of the rast here
                 maxBytes = 10 * 1024 *1024) %>%
  #add a legend
  addRasterLegend(ntems2020, opacity = 1)%>%
  #add scale bar
  addScaleBar(position = c("bottomleft"))


```

```{r map, echo=FALSE}
l
```

<br>
<br>

**Comparing/Showing Multiple Layers:**

Now let's step things up a notch, and add some additional content to our map. For the purposes of this demonstration, let's say you have been tasked with showcasing changes in land cover on Vancouver Island from 2000 to 2020. To do so, we can use 2 key leaflet functionalities: 1) adding a toggle menu - allowing us to switch between layers, 2) adding a 'slider', which we can use to visualize two layers side-by-side.

1\) In order to add a toggle menu and provide NTEMs land cover of both 2000 and 2020 in our map, we do need to provide a name for our layers using the 'group' argument within the 'addRasterImage' function.

```{r leaflet2, include=TRUE, warning = FALSE}


l2 <- leaflet(width = 600, height = 600)%>%
  
  #lets use some base imagery from ESRI
  addProviderTiles("Esri.WorldImagery") %>%
  
  #now lets add our NTEMS land cover from 2000
  addRasterImage(ntems2000, 
                 #you can increase the allowable size of the rast here
                 maxBytes = 10 * 1024 *1024,
                 #provide a name for the layer
                 group = "LC-2000") %>%
  
  #now lets add our NTEMS land cover from 2020
  addRasterImage(ntems2020, 
                 #you can increase the allowable size of the rast here
                 maxBytes = 10 * 1024 *1024,
                 #provide a name for the layer
                 group = "LC-2020") %>%
  
  #allow for layers to be toggles on/off by adding them to the layers control
  addLayersControl(baseGroups = c("LC-2000","LC-2020")) %>%
  
  #add a legend
  addRasterLegend(ntems2000, opacity = 1)%>%
  
  #add scale bar
  addScaleBar(position = c("bottomleft"))


```

```{r map2, echo=FALSE}
l2
```
<br>
<br>

2\) In order to visualize layers side-by-side and compare them using a window slider, we need to create a left and right pane, and assign our NTEMS land cover rasters to one of these panes.

```{r leaflet3, include=TRUE, warning = FALSE}


l3 <- leaflet(width = 600, height = 600)%>%
  
  #set up the two map panes
  addMapPane("right", zIndex = 1) %>%
  addMapPane("left",  zIndex = 2) %>%
  
  #add the ESRI basemap to both map panes
  addProviderTiles("Esri.WorldImagery", group = "base", layerId = "baseid1", options = pathOptions(pane = "right")) %>%
  addProviderTiles("Esri.WorldImagery", group = "base", layerId = "baseid2", options = pathOptions(pane = "left")) %>%
  
  #now lets add our NTEMS land cover from 2000
  addRasterImage(ntems2000, 
                 #you can increase the allowable size of the rast here
                 maxBytes = 10 * 1024 *1024,
                 #provide a name for the layer
                 group = "LC-2000",
                 #assign to a pane
                 options = leafletOptions(pane = "left")) %>%
  
  #now lets add our NTEMS land cover from 2020
  addRasterImage(ntems2020, 
                 #you can increase the allowable size of the rast here
                 maxBytes = 10 * 1024 *1024,
                 #provide a name for the layer
                 group = "LC-2020",
                 #assign to a pane
                 options = leafletOptions(pane = "right")) %>%
  
  #allow for layers to be toggles on/off by adding them to the layers control
  addLayersControl(overlayGroups = c("LC-2000", "LC-2020")) %>%
  
  #add slider control
  addSidebyside(layerId = "sidecontrols",
                rightId = "baseid1",
                leftId  = "baseid2",
                options = list(padding = 0)) %>%
  
  #add a legend
  addRasterLegend(ntems2000, opacity = 1)%>%
  
  #add scale bar
  addScaleBar(position = c("bottomleft"))


```

```{r map3, echo=FALSE}
l3
```
